---
title: "Homework 1"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1
## This problem focuses on variable types and coercion.

First I create a data frame comprised of:


A random sample of size 10 from a uniform[0, 5] distribution

A logical vector indicating whether elements of the sample are greater than 2

A (length-10) character vector

A (length-10) factor vector

```{r}
types_df = tibble(
  vec_unif = runif(10, min = 0, max = 5),
  vec_logical = vec_unif > 2,
  vec_char = c("My", "name", "is", "Yueming", "Kou", "and", "I", "love", "Data", 
               "Science"),
  vec_factor = factor(c("white", "black", "red", "orange", "yellow", "green", "blue", 
                        "purple", "grey", "pink"))
)
```

The chunk below try to take the mean of each variable in types_df

```{r}
mean(types_df$vec_unif)
mean(types_df$vec_logical)
mean(types_df$vec_char)
mean(types_df$vec_factor)
```

The "mean()" function works for the random sample and the logical vector,
but it doesn't work for the character vector and the factor vector.
The reason is elements of the character vector and the factor vector are not numeric or logical.

The chunk below applies the "as.numeric" function to the logical, character, and factor variables.

```{r, results = 'hide'}
as.numeric(types_df$vec_unif)
as.numeric(types_df$vec_logical)
as.numeric(types_df$vec_char)
as.numeric(types_df$vec_factor)
```

Notice that NAs introduced by coercion when applying the "as.numeric" function to the character variables.


The chunk below converts the character variable from character to factor to numeric; similarly, converts the factor variable from factor to character to numeric.

```{r}
char_to_factor = as.factor(types_df$vec_char)
char_to_factor
as.numeric(char_to_factor)
factor_to_char = as.character(types_df$vec_factor)
factor_to_char
as.numeric(factor_to_char)
```

Notice that NAs introduced by coercion when converting the factor variable from factor to charcter to numeric.

# Problem 2
## This problem focuses on plotting and the use of inline R code.

First I create a data frame comprised of:


x: a random sample of size 1000 from a standard Normal distribution

y: a random sample of size 1000 from a standard Normal distribution

A logical vector indicating whether the x + y > 0

A numeric vector created by coercing the above logical vector

A factor vector created by coercing the above logical vector

```{r df_create}
set.seed(1)

plot_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  plot_logical = x + y > 0,
  plot_numeric = as.numeric(plot_logical),
  plot_factor = as.factor(plot_logical)
)

```

First I write a short description of my vector using inline R code.

The size of the dataset is `r length(plot_df)`.

The mean of x is `r mean(plot_df$x)`.

The median of x is `r median(plot_df$x)`.



The chunk below makes a scatterplot of y vs x; color points using the logical variable.

```{r}
ggplot(plot_df, aes(x, y, type = "n"))
ggplot(subset(plot_df, plot_logical == 'TRUE'), aes(x, y, col = 'red')) + geom_point()
ggplot(subset(plot_df, plot_logical == 'FALSE'), aes(x, y, col = 'blue')) + geom_point()

with(plot_df, plot(x, y, type = "n")) 
with(subset(plot_df, plot_logical == 'TRUE'), points(x, y, col = 'red'))
with(subset(plot_df, plot_logical == 'FALSE'), points(x, y, col = 'blue'))
legend("bottomright", pch = 1, col = c("red","blue"), legend = c("plot_logical = 'TRUE'", "plot_logical = 'FALSE'"))

```

The chunk below makes a scatterplot of y vs x; color points using the numeric variable.

```{r}
with(plot_df, plot(x, y, type = "n")) 
with(subset(plot_df, plot_numeric == '1'), points(x, y, col = 'red'))
with(subset(plot_df, plot_numeric == '0'), points(x, y, col = 'blue'))
legend("bottomright", pch = 1, col = c("red","blue"), legend = c("plot_numeric = '1'","plot_numeric = '0'"))

```

The chunk below makes a scatterplot of y vs x; color points using the factor variable.

```{r}
with(plot_df, plot(x, y, type = "n")) 
with(subset(plot_df, plot_factor == 'TRUE'), points(x, y, col = 'red'))
with(subset(plot_df, plot_factor == 'FALSE'), points(x, y, col = 'blue'))
legend("bottomright", pch = 1, col = c("red","blue"), legend = c("plot_factor = 'TRUE'","plot_factor = 'FALSE'"))
```
